Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-11-30 demystify.tex
updates

@@

>>>>

\newtcolorbox{myemphasisbox}{
  enhanced,
  boxrule=1pt, frame hidden,
  %borderline = {1pt}{0pt}{blue!75!black},
  colback=gray!20!white,
  sharp corners
}

\newenvironment{emphasisbox}[1][]{
  \small
  \begin{myemphasisbox}
    {\small \textbf{#1}}
  }{
  \end{myemphasisbox}
}

@@

\section{``Chance me'' on getting admitted to Top 10}\label{sec:chance-me}

Many students, especially international students, aim for very top CS PhD programs such as Stanford and MIT, and Ivy Leagues. Every year Reddit and other forums have numerous students asking for evaluation of their chances of getting into these schools (the so called \emph{``chance me''} or \emph{``roast my CV/profile''} posts) and then later posts on being \emph{``rejected everywhere''}.

\paragraph{You are unlikely to get in these schools}  While being ambitious is always good, it is important to be realistic, and the harsh reality is that it is very unlikely that you will get in MIT, or Harvard. Why? people who do get in these schools often were explicitly encouraged by their mentors and LoR writers, who themselves might be alumni of these schools or are academic celebrities who've sent many of their students there. If you are not in this category (doubt you are, because you're asking random people on the internet instead of your trusted LoR writers), then you're likely not going to get in, and MIT and Stanford would glady take your application money.

\paragraph{``Your'' top might not really top in CS} Harvard and the Ivies? yes good for many things but usually not coding marathon and typically MIA at top CS confs (ok, they are still good, but a school somewhere in Maryland, Wisconsin, or Urbana Champaign cornfield would likely beat them by far in CS). MIT and Stanford? yes, they are top, but many are equivalent, and there could be CS areas where they are not as strong as other schools.  Caltech and John Hopkins?  ... are you confused CS with Physics and Medicine?

You're missing out some serious places: UCSD, UW, UMD, Michigan, Purdue, Stony brook. These all have super strong CS programs but rarely get mentioned. Utah, Rutgers, Buffalo, Penn State, UVA, Ohio, UC Riverside, George Mason ... any of them rings a bell? These might not ranked very high on QS or US News and are not household names to you, your friends, mom and dad, but they are excellent CS schools where graduating students from your dream schools often apply for faculty positions (e.g., see \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU}{GMU stats}). You can go to these places and be superstar there.

\paragraph{You don't need to be top}  Top schools will open doors to more opportunities. In countries like Vietnam, there are usually only a few top schools, and if you don't get into them, you are doomed  (or so you think). But that is not the case in the US. There are simply too many good schools, and it is not a binary choice between top X or nothing.  In fact, ``equity'' is emphasized in the US and ``smaller'' schools are often given more resources and attention to help students succeed.

\begin{emphasisbox}
In short, don't aim for just the moon and the stars, also try the sky, clouds, or even just the trees.  You might find much happiness and success there, which then can catapult you places beyond the moon and the stars.  This is the American dream, which is strongly embodied in the its higher education system.
\end{emphasisbox}
